---
title: "ProblemSet2"
author: "Xuejing"
date: "16/11/2021"
output:
  html_document:
    theme: cerulean
    number_sections: no
    toc: yes
    toc_float: yes
    code_folding: hide
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r libraries, include=FALSE}
library(ggthemes)
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(lubridate)
library(extrafont)
library(vroom)
library(ggtext)
library(gapminder)
library(ggrepel)
library(patchwork)
library(gghighlight)
library(skimr)
library(Rcpp)
library(janitor) # clean_names()
library(ggforce)
library(ggmap)
library(gridExtra)
library(pander)
library(here)
library(leaflet)
library(tmap)
library(tmaptools)
library(hrbrthemes)
library(mapview)
library(viridis)
# remotes::install_github("kjhealy/nycdogs")
library(nycdogs)
library(sf) # for geospatial visualisation

loadfonts(device="pdf")
# loadfonts(device="pdf") if you have a mac

```


```{r ICE ,message=FALSE,warning=FALSE}
# Read data
data_dir <- here::here("data/stop-search")

files <- fs::dir_ls(path = data_dir, regexp = "\\.csv$", recurse = TRUE) 
#recurse=TRUE will recursively look for files further down into any folders

#read them all in using vroom::vroom()
stop_search_data <- vroom(files, id = "source")

#cleaning
stop_search_cleaned <- stop_search_data %>% 
  clean_names() %>%
  select(-source, -part_of_a_policing_operation, -policing_operation, - outcome_linked_to_object_of_search,
         -removal_of_more_than_just_outer_clothing)

stop_search_cleaned <- stop_search_cleaned %>% 
  mutate(self_defined_ethnicity = case_when(grepl("^Asian", self_defined_ethnicity) ~ "Asian",
                                            grepl("^Black", self_defined_ethnicity) ~ "Black",
                                            grepl("^White", self_defined_ethnicity) ~ "White",
                                            T ~ "Other"),
         object_of_search = case_when(object_of_search == "Controlled drugs" ~ "drugs",
                                      object_of_search %in% c("Offensive weapns", "Anything to threaten or harm anyone") ~ "weapons",
                                      object_of_search == "Stolen goods" ~ "stolen goods",
                                      T ~ "other"),
         outcome = case_when(outcome == "A no further action disposal" ~ "disposal",
                             outcome  == "Arrest" ~ "arrest",
                             outcome == "Community resolution" ~ "community resolution",
                             T ~ "other"))

stop_search_cleaned <- stop_search_cleaned %>% 
  drop_na() %>% 
  rename(lat = latitude,
         lng = longitude) %>% 
  mutate(age_range = fct_relevel(age_range, c("under 10", "10-17", "18-24", "25-34", "over 34")),
         month = month(date),
         year = year(date))

```
### Plot on London Map 

```{r}
stop_search_sf <-  st_as_sf(stop_search_cleaned, 
                              coords=c('lng', 'lat'), 
                              crs = 4326)
st_geometry(stop_search_sf)

london_wards_sf <- read_sf(here::here("data/London-wards-2018_ESRI","London_Ward.shp"))
st_geometry(london_wards_sf)

# change the CRS to use WGS84 lng/lat pairs
london_wgs84 <-  london_wards_sf %>% 
  st_transform(4326)


#Plot against gender
ggplot() +
  geom_sf(
    data = london_wgs84, fill = "cornsilk", size = 0.1, colour = "royalblue4" ) +
  geom_point(
    data = stop_search_cleaned %>% filter(lat < 51.8 & lng > -0.6 & outcome == "arrest"), aes(x = lng, y = lat, fill = gender), size = 1.5, alpha = 0.7, shape = 21
  ) + 
  scale_fill_manual(values=c("hotpink1","lightblue1","grey"))+
  coord_sf(datum = NA) +
#  facet_grid(officer_defined_ethnicity ~ self_defined_ethnicity, labeller = label_both +
  labs(
    title = "The majority of people arrested in Stop-and-search are male",
    subtitle = "Stop-and-search location in London mapped with target gender",
    x = "",
    y = "",
    fill = "Gender"
  ) +
  theme_minimal() +
  NULL


#Plot against ethnicity
ggplot() +
  geom_sf(
    data = london_wgs84, fill = "cornsilk", size = 0.1, colour = "royalblue4" ) +
  geom_point(
    data = stop_search_cleaned %>% filter(lat < 51.8 & lng > -0.6 & outcome == "arrest"), aes(x = lng, y = lat, fill = officer_defined_ethnicity), size = 1.5, alpha = 0.7, shape = 21
  ) + 
  scale_fill_manual(values=c("tan","firebrick","azure1","snow2"))+
  coord_sf(datum = NA) +
  labs(
    title = "A significant number arrested in Stop-and-search are black people",
    subtitle = "Stop-and-search location in London mapped with target ethnicity",
    x = "",
    y = "",
    fill = "Ethnicity"
  ) +
  theme_minimal() +
  NULL


tmap::tmap_mode("view") # interactive map

```

```{r}
london_wgs84_2 <- london_wgs84 %>%
  mutate(count = lengths(
    st_contains(london_wgs84, 
                stop_search_sf %>%
                  filter(outcome == "arrest")))) 


ggplot(data = london_wgs84_2, aes(fill = count)) +
  geom_sf() +
  scale_fill_gradient(low = "snow1", high = "firebrick") +
  theme_minimal() +
  coord_sf(datum = NA) +
  labs(title = "Central london has higher rate of arrest",
       subtitle = "Number of stop-and-search arrests mapped with London wards",
       fill = "Number of Arrests") +
  theme(axis.text = element_blank()) +
  theme(strip.text = element_text(color = "white")) +
  NULL


tmap::tmap_mode("view") # interactive map
tmap::tm_shape(london_wgs84_2) +
  tm_polygons("count")


```
### Thought piece on the visualization

With a significantly larger dataset, it makes sense to analyze the stop-and-search of which the outcomes are _Arrest_. In the three map plots, I have shown that the majority of people arrested in stop-and-searches are male, and a significant amount of arrests happen on black people. Central London has a higher density in terms of number of arrests.

When designing the plots, the CRAP principle was adopted as colors I chose has high contrast, especially for the part I want to emphasize on. Titles and subtitles are top-left aligned. 

Evaluated based on Alberto Cairo’s 5 qualities for a great visualization, the visulizations conveys the right message without manipulating axis to mislead the audience. They are functional in telling the story, aesthetically pleasing in design and color, insightful as they mapped the data with geographical information, and enlightening as they reveal the distribution of stop-and-search in London.
